---
title: "Lab2"
format:
  html: 
   code-fold: true
   message: false
editor: visual
---

```{r}
#| echo: false 
#| message: false
here::i_am("lab2.qmd")
library(here)
```

### Exercise 1

```{r}
spotify <- readRDS("top-spotify-songs-from-20102019-by-year.Rds")
library(dplyr)

```

## Question 1

```{r}
resultsong <- spotify %>% distinct(title)
```

The data set contains `r nrow(resultsong)` songs.

```{r}
resultartist <- spotify %>% distinct(`artist`)
```

The data set contains `r nrow(resultartist)` artists.

```{r}
resultgenre <- spotify %>% distinct(`top genre`)
```

The data set contains `r nrow(resultgenre)` musical genre.

## Question 2

```{r}
songsperyear <- spotify |>
  group_by(year) |>
  summarize("number of songs" = n())
knitr::kable(songsperyear)
```

## Question 3

```{r}
artistsong <- spotify %>% distinct(`artist`, title)
mostpopular <-artistsong |>
  group_by(`artist`) |>
  summarize(number_of_songs = n()) |>
  arrange(number_of_songs) |>
  slice_max(number_of_songs)
```

The artist with the largest number of songs in the data set is `r mostpopular[1]` and the number of song that the artist made is `r mostpopular[,2]`.

## Question 4

```{r}
resultgenre <- spotify |>
    group_by(`top genre`)|>
  summarise("mean tempo"=mean(bpm),
          "median tempo"=median(bpm),
          "minimum tempo"= min(bpm),
          "maximum tempo"= max(bpm),
         " number of songs"=n()
         )
knitr::kable(resultgenre)
```

## Question 5

```{r}
resultperyear <- spotify %>% distinct(spotify$title, .keep_all = TRUE) |>
  group_by(year) |>
  summarise(
    median_energy = median(nrgy),
median_danceability= median(dnce)) 

```

## Question 6

```{r}
library(ggplot2)
library(tidyr)
```

```{r}
long_resultperyear <- resultperyear|> pivot_longer(-year, names_to = "median", values_to="Value")
ggplot(long_resultperyear, aes(x=year,y=Value, group=`median`, color=`median`)) +
  geom_line() +
  scale_x_continuous(breaks = seq(min(resultperyear$year), max(resultperyear$year), by = 2)) + 
  labs(title = "Temporal evolution of median annual energy and danceability",
       x = "Year",
       y = "Median Value") 
```

### Exercise 2

```{r}
dropout <- readRDS("dropout.Rds")
library(dplyr)
```

## Question 1

```{r}
q1 <- dropout |>
  group_by(Gender,`Marital status`) |>
  summarize(median_age_at_enrollment=median(`Age at enrollment`))
q1[, 3] <- lapply(q1[, 3], function(x) round(x))
```

## Question 2

```{r}
q2 <- q1 %>%
  pivot_wider(names_from = Gender, values_from = median_age_at_enrollment)
knitr::kable(q2)

```

## Question 3

```{r}
q3 <- dropout |>
  select(Target, starts_with("Curricular units 1st sem (credited)"):ends_with("Curricular units 2nd sem (without evaluations)")) |>
  group_by(Target) |>
  summarize(
    across(starts_with("Curricular units 1st sem (credited)"):ends_with("Curricular units 2nd sem (without evaluations)"), ~ mean(.), .names = "mean {.col}")
  )
```

## Question 4

```{r}
q4 <- q3 |>
  pivot_longer(
    cols = -Target, names_to = "Units", values_to = "Mean") |>
  pivot_wider(names_from = Target,values_from = Mean )

q4[, 2:4] <- lapply(q4[, 2:4], function(x) round(x, 2))

knitr::kable(q4)

```
